Template for indicator documentation

### 2023-09-12 DRAFT_insectIndicators.Rmd

# Notes on this draft
Most things are done but some things needs updates or touching up.

@@

* Upgrade genetic diversity calculation. The current calculation is just a placeholder, awaiting some new data from Marie. Waiting for software installation by IT. Unsure of timeline.
* Calculate NorIns indicators on 5 year interval. (Moving window?)
* Fix remaining comments in this draft inside hard brackets.

@@

```{r}
getDiffWeights()

```
```{r}
getAmountWeights()
```
Lastly, we specify the number of samples for the bootstrap. To speed up, only 999 samples are used here.

@@

```{r}
hInd2022TrondGress <- calcInd(type = "Humler",
                              region_short = "Trond",
                              habitat = "Gressmark",
                              year = 2022,
                              nIter = nIter,
                              save.draws = T)
```

@@

```{r}
connect_to_insect_db()
norway_regions <- Norimon::get_map()
norway_reg_NorIns <- norway_regions %>% 
  filter(region != "Vestlandet") %>% 
  select(region) %>% 
  group_by(region) %>% 
  summarize(geom = st_union(geom))
```

@@

Fetch the data.
```{r}
biomass_sn <- get_biomass(subset_year = 2020:2022,
                          subset_region = NULL)
```
Calculate the indicator values for each region and year.
```{r}
biomass_sn_boot <- bootstrap_value(df = biomass_sn,
                                   value = avg_wet_weight,
                                   groups = c("year",
                                              "region_name")
                                   )
```

@@

Fetch the data.
```{r}
richness_sn <- get_observations(subset_year = 2020:2022,
                                subset_region = NULL)
```
Calculate the indicator values for each region and year.
```{r}
richness_sn_boot <- bootstrap_value(df = richness_sn,
                                   value = no_species,
                                   groups = c("year",
                                              "region_name")
                                   )
```

@@

pollinators_sn <- get_observations(subset_year = 2020:2022,
                                   subset_families = pollinator_fam,
                                   subset_region = NULL)
```
Calculate the indicator values for each region and year.
```{r}
pollinators_sn_boot <- bootstrap_value(df = pollinators_sn,
                                       value = no_species,
                                       groups = c("year",
                                                  "region_name")
                                       )
```
Compare to arbitrary reference value and normalize.
```{r}
pollinators_ref <- 50

@@

### Bumblebees in open landscapes
```{r}
require(bombLepiSurv)
humlesommerfConnect(host = "ninradardata01.nina.no")

@@

```{r bombusGressBarplot, fig.cap = "Bumblebees in grasslands 2021"}
allBombusGressmark2022 <- getAllData(type = "bumblebees",
                          habitat = "gressmark",
                          year = 2022,
                          language = "norsk")

surveyBarplot(allBombusGressmark2022)

@@

```{r}
nIter = 999

@@

```{r}
hInd2022SorGress <- calcInd(type = "Humler",
                              region_short = "Sor",
                              habitat = "Gressmark",
                              year = 2022,
                              nIter = nIter,
                              save.draws = T)
```

```{r}
hInd2022OstGress <- calcInd(type = "Humler",
                              region_short = "Ost",
                              habitat = "Gressmark",
                              year = 2022,
                              nIter = nIter,
                              save.draws = T)
```

```{r}
hInd2022VestGress <- calcInd(type = "Humler",
                              region_short = "Vest",
                              habitat = "Gressmark",
                              year = 2022,
                              nIter = nIter,
                              save.draws = T)
```

@@

```{r}
nIter = 999

@@

```{r}
sInd2022SorGress <- calcInd(type = "Sommerfugl",
                              region_short = "Sor",
                              habitat = "Gressmark",
                              year = 2022,
                              nIter = nIter,
                              save.draws = T)
```

```{r}
sInd2022OstGress <- calcInd(type = "Sommerfugl",
                              region_short = "Ost",
                              habitat = "Gressmark",
                              year = 2022,
                              nIter = nIter,
                              save.draws = T)
```

```{r}
sInd2022VestGress <- calcInd(type = "Sommerfugl",
                              region_short = "Vest",
                              habitat = "Gressmark",
                              year = 2022,
                              nIter = nIter,
                              save.draws = T)
```

>>>>

<!--# Notes on this draft
Most things are done but some things needs updates or touching up.

@@

* Upgrade genetic diversity calculation. The current calculation is just a placeholder, awaiting some new data from Marie. Waiting for software installation by IT. Unsure of timeline.
* Calculate NorIns indicators on 5 year interval. (Moving window?)
* Fix remaining comments in this draft inside hard brackets.-->

@@

```{r, eval=FALSE}
getDiffWeights()

```{r, echo=FALSE}
diffWeights<-readRDS("data/cache/diffWeights.RDS")
diffWeights
```

```{r, eval=FALSE}
getAmountWeights()
```

```{r, echo=FALSE}
amountWeights<-readRDS("data/cache/amountWeights.RDS")
amountWeights
```

Lastly, we specify the number of samples for the bootstrap. To speed up, only 999 samples are used here.

@@

```{r, eval=FALSE}
hInd2022TrondGress <- calcInd(type = "Humler",
                              region_short = "Trond",
                              habitat = "Gressmark",
                              year = 2022,
                              nIter = nIter,
                              save.draws = T)
```

```{r, echo=FALSE}
hInd2022TrondGress<-readRDS("data/cache/hInd2022TrondGress.RDS")
```

@@

```{r}
#connect_to_insect_db()
norway_regions <- Norimon::get_map()
norway_reg_NorIns <- norway_regions %>% 
  filter(region != "Vestlandet") %>% 
  select(region) %>% 
  group_by(region) %>% 
  summarize(geom = st_union(geom))
```

@@

Fetch the data.
```{r, eval=FALSE}
biomass_sn <- get_biomass(subset_year = 2020:2022,
                          subset_region = NULL)
```

```{r, echo=FALSE}
biomass_sn<-readRDS("data/cache/biomass_sn.RDS")
```

Calculate the indicator values for each region and year.
```{r}
biomass_sn_boot <- bootstrap_value(df = biomass_sn,
                                   value = avg_wet_weight,
                                   groups = c("year",
                                              "region_name")
                                   )
```

@@

Fetch the data.
```{r, eval=FALSE}
richness_sn <- get_observations(subset_year = 2020:2022,
                                subset_region = NULL)
```

```{r, echo=FALSE}
richness_sn<-readRDS("data/cache/richness_sn.RDS")
```

Calculate the indicator values for each region and year.
```{r}
richness_sn_boot <- bootstrap_value(df = richness_sn,
                                   value = no_species,
                                   groups = c("year",
                                              "region_name")
                                   )
```

@@

```{r, eval=FALSE}
pollinators_sn <- get_observations(subset_year = 2020:2022,
                                   subset_families = pollinator_fam,
                                   subset_region = NULL)
```

```{r, echo=FALSE}
pollinators_sn<-readRDS("data/cache(pollinators_sn.RDS")
```

Calculate the indicator values for each region and year.
```{r}
pollinators_sn_boot <- bootstrap_value(df = pollinators_sn,
                                       value = no_species,
                                       groups = c("year",
                                                  "region_name")
                                       )
```
Compare to arbitrary reference value and normalize.
```{r}
pollinators_ref <- 50

@@

### Bumblebees in open landscapes
```{r, eval=FALSE}
require(bombLepiSurv)
humlesommerfConnect(host = "ninradardata01.nina.no")

@@

```{r bombusGressBarplot,eval=FALSE, fig.cap = "Bumblebees in grasslands 2021"}
allBombusGressmark2022 <- getAllData(type = "bumblebees",
                          habitat = "gressmark",
                          year = 2022,
                          language = "norsk")

@@

```{r, echo=FALSE}
allBombusGressmark2022<-readRDS("data/cache/allBombusGressmark2022.RDS")

surveyBarplot(allBombusGressmark2022)
```

```{r, eval=FALSE}
nIter = 999

@@

```{r, eval=FALSE}
hInd2022SorGress <- calcInd(type = "Humler",
                              region_short = "Sor",
                              habitat = "Gressmark",
                              year = 2022,
                              nIter = nIter,
                              save.draws = T)
```

```{r, eval=FALSE}
hInd2022OstGress <- calcInd(type = "Humler",
                              region_short = "Ost",
                              habitat = "Gressmark",
                              year = 2022,
                              nIter = nIter,
                              save.draws = T)
```

```{r, eval=FALSE}
hInd2022VestGress <- calcInd(type = "Humler",
                              region_short = "Vest",
                              habitat = "Gressmark",
                              year = 2022,
                              nIter = nIter,
                              save.draws = T)
```

@@

```{r, eval=FALSE}
nIter = 999

@@

```{r, eval=FALSE}
sInd2022SorGress <- calcInd(type = "Sommerfugl",
                              region_short = "Sor",
                              habitat = "Gressmark",
                              year = 2022,
                              nIter = nIter,
                              save.draws = T)
```

```{r, eval=FALSE}
sInd2022OstGress <- calcInd(type = "Sommerfugl",
                              region_short = "Ost",
                              habitat = "Gressmark",
                              year = 2022,
                              nIter = nIter,
                              save.draws = T)
```

```{r, eval=FALSE}
sInd2022VestGress <- calcInd(type = "Sommerfugl",
                              region_short = "Vest",
                              habitat = "Gressmark",
                              year = 2022,
                              nIter = nIter,
                              save.draws = T)
```
